---
title: "NLP and Topic Modelling"
author: "Alta de Waal"
date: "21 February 2019"
output: html_document
---

#Prepwork

## Load the libraries

```{r message=FALSE, warning=FALSE}
#most of the libraries needed
library(tidytext) #text mining, unnesting
library(topicmodels) #the LDA algorithm
library(tidyr) #gather()
library(dplyr) #awesome tools
library(ggplot2) #visualization
library(kableExtra) #create attractive tables
library(knitr) #simple table generator
library(ggrepel) #text and label geoms for ggplot2
library(gridExtra)
library(formattable) #color tile and color bar in `kables`
library(tm) #text mining
library(circlize) #already loaded, but just being comprehensive
library(plotly) #interactive ggplot graphs
```

## Functions
```{r echo=FALSE}
#define some colors to use throughout
my_colors <- c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#D55E00", "#D65E00")

#customize ggplot2's default theme settings
#this tutorial doesn't actually pass any parameters, but you may use it again in future tutorials so it's nice to have the options
theme_lyrics <- function(aticks = element_blank(),
                         pgminor = element_blank(),
                         lt = element_blank(),
                         lp = "none")
{
  theme(plot.title = element_text(hjust = 0.5), #center the title
        axis.ticks = aticks, #set axis ticks to on or off
        panel.grid.minor = pgminor, #turn on or off the minor grid lines
        legend.title = lt, #turn on or off the legend title
        legend.position = lp) #turn on or off the legend
}

#customize the text tables for consistency using HTML formatting
my_kable_styling <- function(dat, caption) {
  kable(dat, "html", escape = FALSE, caption = caption) %>%
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"),
                full_width = FALSE)
}

word_chart <- function(data, input, title) {
  data %>%
  #set y = 1 to just plot one variable and use word as the label
  ggplot(aes(as.factor(row), 1, label = input, fill = factor(topic) )) +
  #you want the words, not the points
  geom_point(color = "transparent") +
  #make sure the labels don't overlap
  geom_label_repel(nudge_x = .2,  
                   direction = "y",
                   box.padding = 0.1,
                   segment.color = "transparent",
                   size = 3) +
  facet_grid(~topic) +
  theme_lyrics() +
  theme(axis.text.y = element_blank(), axis.text.x = element_blank(),
        #axis.title.x = element_text(size = 9),
        panel.grid = element_blank(), panel.background = element_blank(),
        panel.border = element_rect("lightgray", fill = NA),
        strip.text.x = element_text(size = 9)) +
  labs(x = NULL, y = NULL, title = title) +
    #xlab(NULL) + ylab(NULL) +
  #ggtitle(title) +
  coord_flip()
}
```

## Load the data
The three data sources are: Prince lyrics, a book called "Machine Learning For Dummies", and another book called "Why Icebergs Float", each having over 30,000 (non-distinct) words
```{r}
three_sources_tidy_balanced <- read.csv("three_sources_tidy_balanced.csv",
                                        stringsAsFactors = FALSE)
```

# Build the LDA model

## 1. Examine the data
```{r echo=FALSE, message=FALSE, warning=FALSE}

#group the dataset by writer (source) and count the words
three_sources_tidy_balanced %>%
  group_by(source) %>%
  mutate(word_count = n()) %>%
  select(source, genre, word_count) %>% #only need these fields
  distinct() %>%
  ungroup() %>%
  #assign color bar for word_count that varies according to size
  #create static color for source and genre
  mutate(word_count = color_bar("lightpink")(word_count),  
         source = color_tile("lightblue","lightblue")(source),
         genre = color_tile("lightgreen","lightgreen")(genre)) %>%
  my_kable_styling("Three Sources Stats")
```
## 2. Create the Document-Term Matrix
Bag-of-words vectorisation
```{r}
three_sources_dtm_balanced <- three_sources_tidy_balanced %>%
  #get word count per document to pass to cast_dtm
  count(document, word, sort = TRUE) %>%
  ungroup() %>%
  #create a DTM with docs as rows and words as columns
  cast_dtm(document, word, n)
#examine the structure of the DTM
three_sources_dtm_balanced
```

## 3. Set Variables
```{r}
#assign the source dataset to generic var names
#so we can use a generic function per model
source_dtm <- three_sources_dtm_balanced
source_tidy <- three_sources_tidy_balanced
```

## 4. Fit the model

```{r}
k <- 3 #number of topics
seed = 1234 #necessary for reproducibility
#fit the model passing the parameters discussed above
#you could have more control parameters but will just use seed here
lda <- LDA(source_dtm, k = k, method = "GIBBS", control = list(seed = seed))
#examine the class of the LDA object
class(lda)
```

```{r}
#convert the LDA object into a tidy format
#passing "beta" shows the word probabilities
#filter on the word iceberg as an example
#results show probability of iceberg for each topic
tidy(lda, matrix = "beta") %>% filter(term == "iceberg")
```

## 5. Analysing the output
### Identify themes with top words

```{r}
num_words <- 10 #number of words to visualize

#create function that accepts the lda model and num word to display
top_terms_per_topic <- function(lda_model, num_words) {

  #tidy LDA object to get word, topic, and probability (beta)
  topics_tidy <- tidy(lda_model, matrix = "beta")


  top_terms <- topics_tidy %>%
  group_by(topic) %>%
  arrange(topic, desc(beta)) %>%
  #get the top num_words PER topic
  slice(seq_len(num_words)) %>%
  arrange(topic, beta) %>%
  #row is required for the word_chart() function
  mutate(row = row_number()) %>%
  ungroup() %>%
  #add the word Topic to the topic labels
  mutate(topic = paste("Topic", topic, sep = " "))
  #create a title to pass to word_chart
  title <- paste("LDA Top Terms for", k, "Topics")
  #call the word_chart function you built in prep work
  word_chart(top_terms, top_terms$term, title)
}
#call the function you just built!
top_terms_per_topic(lda, num_words)
```
### Classify documents

```{r}
#this time use gamma to look at the prob a doc is in a topic
#just look at the Prince song 1999 as an example
tidy(lda, matrix = "gamma") %>% filter(document == "1999")
```
### Chord diagram
```{r}
#using tidy with gamma gets document probabilities into topic
#but you only have document, topic and gamma
source_topic_relationship <- tidy(lda, matrix = "gamma") %>%
  #join to orig tidy data by doc to get the source field
  inner_join(three_sources_tidy_balanced, by = "document") %>%
  select(source, topic, gamma) %>%
  group_by(source, topic) %>%
  #get the avg doc gamma value per source/topic
  mutate(mean = mean(gamma)) %>%
  #remove the gamma value as you only need the mean
  select(-gamma) %>%
  #removing gamma created duplicates so remove them
  distinct()

#relabel topics to include the word Topic
source_topic_relationship$topic = paste("Topic", source_topic_relationship$topic, sep = " ")

circos.clear() #very important! Reset the circular layout parameters
#assign colors to the outside bars around the circle
grid.col = c("prince" = my_colors[1],
             "icebergs" = my_colors[2],
             "machine_learning" = my_colors[3],
             "Topic 1" = "grey", "Topic 2" = "grey", "Topic 3" = "grey")

# set the global parameters for the circular layout. Specifically the gap size (15)
#this also determines that topic goes on top half and source on bottom half
circos.par(gap.after = c(rep(5, length(unique(source_topic_relationship[[1]])) - 1), 15,
                         rep(5, length(unique(source_topic_relationship[[2]])) - 1), 15))
#main function that draws the diagram. transparancy goes from 0-1
chordDiagram(source_topic_relationship, grid.col = grid.col, transparency = .2)
title("Relationship Between Topic and Source")
```

