---
title: "RLadies: Starwars & Plotly"
# output: github_document
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load all required packages
library(fivethirtyeight)
data(package = 'fivethirtyeight')

library(tidyr)
library(dplyr)
library(ggplot2)
library(plotly)
library(knitr)

```
**References**:


## Let's look at the data
```{r load}
# load the data
# data set imported from Git
# na values = imported as spaces
starwars = read.csv('/Users/Gabriella_Camara/Documents/RLadies/data/StarWars.csv',
                    na.strings = c('', ' '))

kable(starwars %>% head(5), title = 'Head of StarWars data')

# Summary of data
summary(starwars)
colSums(is.na(starwars))

# Favourite way to view data
# View(starwars)

# useful commands for quick snapshot
table(starwars$Education)
unique(starwars$Gender)

```

From the above, we can see that the data is pretty messy.  
Let's see how we can clean this up.  
I've done this pretty manually as I'm quite specific about my column name, there are alternatives: melt, unlist & lapply first row as column names etc.  

```{r cleanup}
# drop first row, don't want this to be considered as another level
starwars = starwars[-1, ] 
starwars = lapply(starwars, as.character) # in order to not recognise 'Response' as a factor/ level
# note, this conversts the df to a list; we'll convert it back

names(starwars) = c('RespondentID',
                    'Seen',
                    'Fan',
                    'Seen.Episode1',
                    'Seen.Episode2',
                    'Seen.Episode3',
                    'Seen.Episode4',
                    'Seen.Episode5',
                    'Seen.Episode6',
                    'Rank.Episode1',
                    'Rank.Episode2',
                    'Rank.Episode3',
                    'Rank.Episode4',
                    'Rank.Episode5',
                    'Rank.Episode6',
                    'View.HanSolo',
                    'View.LukeSkywalker',
                    'View.PrincessLeiaOrgana',
                    'View.Anakin Skywalker',
                    'View.ObiWanKenobi',
                    'View.EmperorPalpatine',
                    'View.DarthVader',
                    'View.LandoCalrissian',
                    'View.BobaFett',
                    'View.C-3P0',
                    'View.R2D2',
                    'View.JarJarBinks',
                    'View.PadmeAmidala',
                    'View.Yoda',
                    'FirstToShoot',
                    'Familiar.ExpandedUniverse',
                    'Fan.ExpandedUniverse',
                    'Fan.Franchise',
                    'Gender',
                    'Age',
                    'Income',
                    'Education',
                    'Region')


starwars = lapply(starwars, as.factor) # notice number of factor (2 = Yes & No)
starwars = as.data.frame(starwars)

# Replace movie names with 'Yes' for seen
starwars[4:9] = sapply(starwars[4:9], function(x)
                                 ifelse(is.na(x), NA, 'Yes'))


kable(starwars %>% head(5))
```

That's better, at most I think I can work with this data.  
We'll change it later if need be.  

## Objectives
From our dataset, there are a couple of thinsg we'd like to get sight of.  
Most of our data is categorical, so it seems that bar charts will be our go-to as they best display categories.  

* Get a grip of Plotly as a visualization technique. See [Plotly cookbook](https://plotly-book.cpsievert.me/index.html)  
* Understand our viewership by different categories  
* Understand views & ranks per Episode of movie  
* Understand favourability of characters  

## Plotly overview
**Insert novice opionion & understanding here**

## Viewership
Let's explore those best invested in the Starwars Franchise & a galaxy far-far away ...  

```{r views}
# Viewership
# 1. just by seen
# 2. Add gender (color) 
# 3 layout = stacked
p1 = starwars %>%
     plot_ly(x = ~Seen) %>%
     add_histogram()
p1


# color adds another 'dimension'
# additioanlly, let's consider the lists passed to layouts
p2 = starwars %>%
     plot_ly(x = ~Seen, color = ~Gender) %>%
     add_histogram() %>%
     layout(title = 'Viewership by Gender',
            xaxis = list(tickangle = -45),
            yaxis = list(title = 'Count'))
p2

# let's stack the bars instead
p3 = starwars %>%
     plot_ly(x = ~Seen, color = ~Age) %>%
     add_histogram() %>%
     layout(title = 'Viewership by Age',
            xaxis = list(tickangle = -45),
            yaxis = list(title = 'Count'),
            barmode = 'stack') 
p3

# I'm interested in the distribution by age
# note we're not feeding any y-values
p4 = starwars %>%
     plot_ly(x = ~Age) %>%
     add_histogram()
p4
# an almost (not) normal bell curve/ uniform?
```
Things to note:  
* 1186 observations (total number of rows); we know unique by number of levels  
* Majority have seen the movie  

## Let's explore the opions of those that have seen the movies:
### Which StarWars movies have you seen?
```{r}
# only trust those that have seen the movies; filter data to go forward here
# pipes (makkgrittr - spelling to be corrected)
starwars_view = starwars %>%
                filter(Seen == 'Yes')

# tidy: objective to create correct variable & observations
# tidyr package
episodes_seen = starwars_view %>% 
                select('Seen.Episode1':'Seen.Episode6') %>%
                gather(Episode, Response) %>%
                drop_na()

# count & prop seen
episodes_seen = episodes_seen %>%
                group_by(Episode) %>%
                summarise(count = n()) %>%
                mutate(prop = count/sum(count)*100)
# Not quite what I'm looking for, this is the proportion as is, but doesn't correctly conisder the population of people who've seen the movie

# let's try again
number_seen = length(starwars_view$RespondentID)
episodes_seen = episodes_seen %>%
                group_by(Episode) %>%
                summarise(count = n()) %>%
                mutate(prop = count/number_seen*100)

# let's put this on a graph
# note difference between bars & histogram
episodes_seen %>%
  plot_ly(x = ~Episode, y = ~count, name = 'Count of Viewers') %>%
  add_bars() %>%
  add_trace(y = ~prop, name = 'Proportion of Viewers')
# just to show the ability to add a 'trace'; not mst logical chart

# change orientation
# customize colour of bars with rgb scale (mostly a guessing game for me?)
episodes_seen %>%
  plot_ly(x = ~count, y = ~Episode, name = 'Count of Viewers',
          marker = list(color = 'rgb(26, 118, 255)')) %>%
  add_bars()

# # further enrich df with ranking alongside movie
# stars_rank = starwars_view %>% 
#              select('RespondentID', 
#                     'Rank.Episode1', 'Rank.Episode2', 'Rank.Episode3',
#                     'Rank.Episode4', 'Rank.Episode5', 'Rank.Episode6') %>% 
#              gather(Episode, Rank, -RespondentID)
# 
# stars_view = starwars_view %>% 
#              select('RespondentID', 'Seen.Episode1':'Seen.Episode6') %>%
#              gather(Episode, Response, - RespondentID) 
# 
# stars_full = cbind(stars_view, star_rank)

```

### What's the best StarWars movie?
```{r}
# In order to correctly compare, we can only consider the rankings of viewers that have watched all 6 movies

# count yes for seen columns
# replace_na(starwars_view, list(Seen.Episode1= 'No'))
starwars_view[, 4:9][is.na(starwars_view[, 4:9])] = 'No'
starwars_view$num.Seen = rowSums(starwars_view[ , 4:9] == 'Yes')

starwars_all6 = starwars_view %>%
                filter(num.Seen == 6)


count_seen = length(starwars_all6$RespondentID)

seen_all6 = starwars_all6 %>%
            select('Rank.Episode1':'Rank.Episode6') %>%
            gather(Episode, Rank) %>%
            drop_na()

rank_all6 = seen_all6 %>%
            group_by(Episode, Rank) %>%
            summarise(count = n()) %>%
            mutate(prop = count/count_seen*100)

rank_all6 %>%
  filter(Rank == 1) %>%
  plot_ly(x = ~Episode, y = ~count, color = I("red")
          # text = count, textposition = 'auto'
          ) %>%
  add_bars() %>%
  layout(
    title = 'Which is your favourite StarWars movie?',
    xaxis = list(
      type = 'category',
      title = 'Movie Name'
    )
  )

# informative & interactive
rank_all6 %>%
  plot_ly(x = ~Episode, y = ~count, color = ~Rank) %>%
  add_bars() %>%
  layout(
    title = 'Which is your favourite StarWars movie?'
  )

p1 = ggplot()
ggplotly(p1)
# best vs worst
  
```

### Characters favourability rankings
```{r}
character_rating = starwars_view %>%
                   select('View.HanSolo':'View.Yoda') %>%
                   gather(Character, Rating) %>%
                   drop_na()

character_rating = character_rating %>%
                   group_by(Character, Rating) %>%
                   summarise(count = n())

# Option 1
character_rating %>%
  plot_ly(x=~count, y = ~Character, color = ~Rating) %>%
  add_bars() %>%
  layout(
    barmode = 'stack'
  )

# Option 2
p1 = character_rating %>%
     filter(Rating == c('Very favorably', 'Somewhat favorably')) %>%
     plot_ly(x=~count, y = ~Character) %>%
     add_bars() %>%
     layout(
       title = 'Favorably viewed'
     )

p2 = character_rating %>%
     filter(Rating == 'Somewhat unfavorably') %>%
     plot_ly(x=~count, y = ~Character) %>%
     add_bars() %>%
     layout(
       title = 'Unfavorably viewed'
     )

subplot(p1, p2) %>% hide_legend()
```

## Now your turn
* View the first to shoot?  
* Demographics of fans?  

